On The Impact of Atoms of Confusion in JavaScript Code

### 2021-09-24 0_Abstract.tex
Updates from Overleaf

\begin{abstract}
{\bf Context.} Engineering maintainable software is essential and must take into account \emph{how understandable the code is} from a human perspective, and spending much cognitive effort might discourage a developer when trying to familiarize herself with a legacy codebase. 
\noindent{\bf Problem.} 
%Besides lack of experience and other factors, 
Previous research works have discussed the impact of small code patterns that contribute to misunderstanding the source code written in statically typed languages (e.g., \clang, \cpplang, and Swift). However, little is known about the influence of these small code patterns (hereafter atoms of confusion) on understanding dynamically typed languages. 
\noindent{\bf Research Method and Goal.} In this work, we fill this research gap by presenting the results of a mixed-methods research:
a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort. Our goal is to investigate the impact of atoms of confusion on understanding JavaScript code---a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development).
\noindent{\bf Results.} From our survey and interviews, we confirm that atoms of confusion lead to a code that is hard to understand: 
%removing the atoms of confusion lead to an improvement above 15\% in 6 out of 9 atoms we consider in our study when developers try to predict the outcomes of code examples. 
developers correctly predict the outcome of at least 15\% more when an atom of confusion is not present for 6 out of 9 atoms.
Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the \texttt{comma operator} is present in 11\% of the projects in our dataset; while five other atoms appear in more than 60\% of the projects---including the atoms \texttt{assignment as value} and \texttt{omitted curly braces}, which appear \num{8253} and \num{53357} times, respectively. Our investigation also reveals additional JavaScript constructs and idioms that might hinder developers from understanding JavaScript code. For instance, even experienced JavaScript developers agree that it is hard to understand code that uses callback functions, even though we found widespread use of \texttt{callbacks} in our projects dataset.  
\noindent{\bf Implications.} Altogether, our findings could help developers write cleaner JavaScript code and contribute to the design of new program transformation tools that aim to simplify JavaScript code. 
\end{abstract}

>>>>

\begin{abstract}
{\bf Context.} Engineering maintainable software is essential and must take into account \emph{how understandable the code is} from a human perspective, and spending much cognitive effort might discourage a developer when trying to familiarize herself with a legacy codebase. 
\noindent{\bf Problem.} 
%Besides lack of experience and other factors, 
Previous research work has discussed the impact of small code patterns that contribute to misunderstanding the source code written in statically typed languages (e.g., \clang, \cpplang, and Swift). However, little is known about the influence of these small code patterns (hereafter atoms of confusion) on understanding dynamically typed languages. 
\noindent{\bf Research Method and Goal.} In this work, we fill this research gap by presenting the results of a mixed-methods research:
a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort. Our goal is to investigate the impact of atoms of confusion on understanding JavaScript code---a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development).
\noindent{\bf Results.} From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: 
%removing the atoms of confusion lead to an improvement above 15\% in 6 out of 9 atoms we consider in our study when developers try to predict the outcomes of code examples. 
developers correctly predict the outcome of at least 15\% more when an atom of confusion is not present for 9 out of 10 atoms. For 5 of these atoms, the difference is statistically significant (accounting for p-value correction) and effect size is large for 4 and medium for 1 of them. 
Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the \texttt{comma operator} is present in 11\% of the projects in our dataset; while five other atoms appear in more than 60\% of the projects---including the atoms \texttt{assignment as value} and \texttt{omitted curly braces}, which appear \num{8253} and \num{53357} times, respectively. Our investigation also reveals additional JavaScript constructs and idioms that might hinder developers from understanding JavaScript code. For instance, even experienced JavaScript developers agree that it is hard to understand code that uses callback functions, even though we found widespread use of \texttt{callbacks} in our projects dataset.  
\noindent{\bf Implications.} Altogether, our findings could help developers write cleaner JavaScript code and contribute to the design of new program transformation tools that aim to simplify JavaScript code. 
\end{abstract}

### 2021-09-24 1_Introduction.tex
Updates from Overleaf

In this way, to better understand the impact of atoms of confusion in JavaScript, in this paper we present a mixed-method research effort based on a survey, on interviews, and on a mining software repositories study. First, we surveyed 140 JavaScript developers. In the survey, their task was to predict the output of small code snippets, with half of them containing atoms of confusion. The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate). The presence of an atom candidate might also affect the time necessary for submitting a correct answer. For instance, the presence of the Comma Operator atom increases the average time to predict the output by 76.23\%, though the outputs of snippets containing the Pre-Increment atom were correctly predicted 38.19\% faster.
To learn more about the misunderstandings in the snippets used in the survey, we also interviewed 15 experienced professional developers. Participants prefered the code snippets without the atom candidates in 70\% of the cases, therefore corroborating the survey findings. The participants also discussed other idioms and constructs that might lead developers to misunderstand JavaScript code.

The frequency in which the atoms appear \emph{in the wild} might reveal cases where developers should be more or less careful when writing JavaScript code. To understand the incidence of atoms, we mined popular JavaScript projects from GitHub and found that atoms of confusion frequently appear in practice. Indeed, four atom candidates are highly used in JavaScript code---the Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.

>>>>

In this way, to better understand the impact of atoms of confusion in JavaScript, in this paper we present a mixed-method research effort based on a survey, on interviews, and on a mining software repositories study. First, we surveyed 140 JavaScript developers. In the survey, their task was to predict the output of small code snippets, with half of them containing atoms of confusion. The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate). For 5 of these atoms, the difference is statistically significant (accounting for p-value correction). Effect size is large for 4 atoms and medium for 1 of them. 
The presence of an atom candidate might also affect the time necessary for submitting a correct answer. For instance, the presence of the Comma Operator atom increases the average time to predict the output by 76.23\%, though the outputs of snippets containing the Pre-Increment atom were correctly predicted 38.19\% faster.
To learn more about the misunderstandings in the snippets used in the survey, we also interviewed 15 experienced professional developers. Participants prefered the code snippets without the atom candidates in 70\% of the cases, therefore corroborating the survey findings. The participants also discussed other idioms and constructs that might lead developers to misunderstand JavaScript code.

The frequency in which the atoms appear \emph{in the wild} might reveal cases where developers should be more or less careful when writing JavaScript code. To understand the incidence of atoms, we mined popular JavaScript projects from GitHub and found that atoms frequently appear in practice. 
Our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems. Four atom candidates appear in 90\% of the analyzed projects and two of them, Implicit Predicate and Ternary Operator, occur more than once for every 100 lines of code in the dataset. 
%Indeed, four atom candidates are frequently used in JavaScript code---the Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.

### 2021-09-24 6_Conclusion.tex
Updates from Overleaf

This paper has reported the results of a mixed-method research effort to better understand the impact of atoms of confusion in JavaScript code. First, we conducted a survey with 140 JavaScript developers. We asked them to predict the output of some code snippets (with and without atoms). We provide evidence that seven atom candidates have a significant impact on program comprehension activities.

Afterwards, we interviewed 15 professional developers. In general, they considered code without atoms of confusion easier to understand and highlighted other JavaScript constructs and idioms that might also introduce misunderstanding. 
%Last, but not least,
Finally, we conducted a mining software repositories study. In particular, we found four atoms of confusion frequently used in JavaScript code, i.e., Implicit Predicate, Ternary Operator, Omitted Curly Braces, and Post Increment.

Our efforts have important implications for practice. First, we alert developers to avoid writing JavaScript code with certain atoms of confusion (e.g., Comma Operator, Automatic Semicolon Insertion, Post Increment, and Omitted Curly Braces). Second, our results might help tool developers to create program transformation tools to remove atoms of confusion automatically. 
%They might firstly focus, for instance, on the atoms we found to be more confusing and at the same time more common in practice (e.g., Omitted Curly Braces and Post Increment).
Promising candidates may be the atoms we found to be most confusing and, at the same time, common in practice (e.g., Omitted Curly Braces and Post Increment).

%As mentioned, the results of the interviews
The results of the interviews pointed out additional JavaScript constructs and idioms
that might introduce misunderstanding. These include not only nested callbacks and prototype-based
inheritance (which might not have a counterpart version), but also idioms such
as \emph{property access} and \emph{arrow functions}. As future work, we intend to
reproduce our survey to validate whether or not these additional
\emph{atom candidates}, which the interviewees pointed out, truly affect
the understanding of JavaScript code.

>>>>

This paper reports the results of a mixed-method research effort to better understand the impact of atoms of confusion in JavaScript code. First, we conducted a survey with 140 JavaScript developers, asking them to predict the output of some code snippets (with and without atoms). We provide evidence that five atom candidates significantly impact program comprehension activities and can be considered atoms of confusion: Comma Operator, Automatic Semicolon Insertion, Post Increment, Assignment as Value, and Implicit Predicate. 
We also interviewed 15 professional developers. In general, they considered code without atom candidates easier to understand and highlighted other JavaScript constructs and idioms that might also introduce misunderstanding. 
%Last, but not least,
Finally, we conducted a mining software repositories study. We found four atom candidates frequently used in JavaScript code, i.e., Implicit Predicate, Post Increment, Ternary Operator, and Omitted Curly Braces.  %The first two can be considered atoms of confusion.

Our efforts have important implications for practice. First, we alert developers to avoid writing JavaScript code with certain atoms of confusion (e.g., Comma Operator, Automatic Semicolon Insertion, Post Increment, Assignment as Value, and Implicit Predicate). Second, our results might help tool developers to create program transformation tools to remove the atoms automatically. 
%They might firstly focus, for instance, on the atoms we found to be more confusing and at the same time more common in practice (e.g., Omitted Curly Braces and Post Increment).
%Promising candidates may be the atoms we found to be most confusing and, at the same time, common in practice (e.g., Implicit Predicate and Post Increment).

%As mentioned, the results of the interviews
The results of the interviews pointed out additional JavaScript constructs and idioms 
that might introduce misunderstanding. These include nested callbacks, % and prototype-based inheritance (which might not have a counterpart version), but also idioms such
%as 
\emph{property access}, and \emph{arrow functions}. As future work, we intend to
reproduce our survey to validate whether or not these additional
\emph{atom candidates}
%, which the interviewees pointed out, 
truly affect
the understanding of JavaScript code.

### 2021-09-24 discussion.tex
Updates from Overleaf

Our work leads to several implications.
First, we generalize the
work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17},
which investigates
the impact of atom candidates on
understanding \clang and \cpplang code. That is,
similarly to their work, the atom candidates
Comma Operator, Post/Pre Increment, Omitted Curly Braces,
Assignment as Value, Implicit Predicate, Logic as
Control Flow, and Ternary Operator make the 
JavaScript code hard to understand. Our results also 
refute the hypothesis that Arithmetic as Logic is an atom of
confusion (i.e., a source of misunderstanding).
In comparison to the original
work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, 
our study
led to some differences in the effect size
of the atom candidates.
Altogether, we answer our first research question
\emph{\rqa}:

\begin{mh}
  {\bf Answer to RQ1:} The first study (survey) gives evidence that
  the atom candidates for \clang and \cpplang
  that also exist for JavaScript correspond to
  a source of misunderstanding in
  JavaScript code. 
\end{mh}

The results of the interview study complement the understanding
of atoms of confusion because the participants make
clear the existence of a trade-off between code comprehension
and other quality attributes. For instance, most of the participants
prefer the version of the code with the Ternary Operator,
even though they agree that its use might contribute to the
misunderstanding of JavaScript code, particularly when novices are maintaining
the codebase. The participants of the interview study also
mentioned other possible sources of misunderstanding in JavaScript,
including the use of prototype-inheritance and nested call-backs (as discussed in Section~\ref{sec:interview-results}). Other JavaScript atom candidates include
Object Destructuring, Array Spread, Object Spread, and Type Conversion.
In summary, the results of the second study (interviews) allow
us to answer the second (\rqb) and third research questions
(\emph{\rqc}).

@@

\begin{mh}
  {\bf Answer to RQ3:} The qualitative analysis of the
  interviews suggest that specific constructs of
  JavaScript might also correspond to atoms of
  confusion, including prototype inheritance and
  nested callbacks. 
\end{mh}

The results of the third study (mining open source
Java script repositories) give evidence that,
although atoms of confusion compromise program
comprehension, they frequently appear in open
source JavaScript projects. In particular,
seven, out of 10 atoms considered
  in our study, appear in more than 50\% of
the projects. 
In summary, the third study
allows us to answer the fourth research
question (\emph{\rqd}).

>>>>

Our work has several implications.
First, it adds external validity to the
work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17},
which investigates
the impact of atom candidates on
understanding \clang and \cpplang code. That is,
similarly to their work, the atom candidates
Comma Operator, Post/Pre Increment, Omitted Curly Braces,
Assignment as Value, Implicit Predicate, Logic as
Control Flow, and Ternary Operator seem to make 
JavaScript code harder to understand. For five of them, the difference is statistically significant, with a large effect size for four atoms. Our results also 
refute the hypothesis that Arithmetic as Logic is an atom of confusion (i.e., a source of misunderstanding).
In comparison to the original
work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, 
our study
led to some differences in the effect size
of the atom candidates.
Altogether, we answer our first research question
\emph{\rqa}:

\begin{mh}
  {\bf Answer to RQ1:} The first study (survey) gives evidence that some of the atom candidates in \clang and \cpplang programs that also exist in JavaScript correspond to a source of misunderstanding in
  JavaScript code. 
\end{mh}

The results of the interview study complement the understanding of atoms of confusion because the participants make clear the existence of a trade-off between code comprehension and other quality attributes. For instance, most of the participants prefer the version of the code with the Ternary Operator, even though they agree that its use might contribute to the misunderstanding of JavaScript code, particularly when novices are maintaining the codebase. The participants of the interview study also
mentioned other possible sources of misunderstanding in JavaScript,
including the use of prototype-based inheritance and nested call-backs (as discussed in Section~\ref{sec:interview-results}). Other JavaScript atom candidates include
Object Destructuring, Array Spread, Object Spread, and Type Conversion.
In summary, the results of the second study (interviews) allow
us to answer the second (\rqb) and third research questions
(\emph{\rqc}).

@@

\begin{mh}
  {\bf Answer to RQ3:} The qualitative analysis of the
  interviews suggests that specific JavaScript constructs might also correspond to atoms of
  confusion, including prototype inheritance and
  nested callbacks. 
\end{mh}

The results of the third study (mining open source
Java script repositories) provides evidence that,
although atoms of confusion compromise program
comprehension, they frequently appear in open
source JavaScript projects. In particular,
seven, out of 10 atoms considered
  in our study, appear in more than 50\% of
the 72 projects we analyzed. Furthermore, at least two of them are used intensively, more than once for every 200 lines of code. In summary, the third study
allows us to answer the fourth research
question (\emph{\rqd}).

### 2021-09-24 main.tex
Updates from Overleaf

\begin{abstract}
  Evolving legacy code is a challenging task, particularly when the code has been poorly written, increasing effort and imposing a significant cognitive load on developers. Although previous research has investigated the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in \clang, \cpplang, and Java, little is known about the influence of these atoms on understanding dynamic languages (such as JavaScript). In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study. From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present inseven out of ten atom candidates. In addition, our MSR effort reveals that atoms of confusion frequently arise in 72 popular open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively. Altogether, our findings might help practitioners (a) better understand the implications of atoms of confusion on understanding JavaScript code, (b) avoid writing code that is unnecessarily difficult to maintain, and (c) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.

>>>>

\begin{abstract}
  Evolving legacy code is a challenging task, particularly when the code has been poorly written, increasing effort and imposing a significant cognitive load on developers. Although previous research has investigated the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in \clang, \cpplang, and Java, little is known about the influence of these atoms on understanding dynamic languages (such as JavaScript). In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study. From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: considering 10 atom candidates, developers correctly predict the outcome of at least 15\% more cases in code snippets where atom candidates are not present, compared to alternative versions that include the atom candidates. For five of these atoms, the difference is statistically significant (accounting for p-value correction). Effect size is large for four atoms and medium for one of them. In addition, our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems. Four atom candidates appear in 90\% of the analyzed projects and two of them occur more than once for every 100 lines of code in the dataset. 
  %: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
  Altogether, our findings might help practitioners (a) better understand the implications of atoms of confusion on understanding JavaScript code, (b) avoid writing code that is unnecessarily difficult to maintain, and (c) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.
